---
title: "Assignment_2_3102_bv10"
author: Bhuvaneswari(Bhu) Venkatraman
netid: bv10
seedid: 3102
output: html_document
---

#### netid: bv10

#### seedid: 3102

# Setting Seed and defining static variables


```{r}
library(glmnet) 
library(pls)
set.seed(3102)
```


# data read, training and test set generation as per dataset file name

```{r}
get_data <- function(dataset_file_name)  {
  
  # load csv
  myData = read.csv(dataset_file_name)
  myData = myData[, -1]
  X = data.matrix(myData[,-1])  
  Y = myData[,1] 
  
  # # generate train and test
  T = 50
  n = length(Y)
  ntest = round(n * 0.25)  # test set size
  ntrain = n - ntest  # training set size
  all.test.id = matrix(0, ntest, T)  #
  for(t in 1:T){
    all.test.id[, t] = sample(1:n, ntest)
  }

  #return necessary inputs to generate prediction
  return(list(myData=myData, X=X, Y=Y, all.test.id=all.test.id))
}
```

# full model prediction and MSE calculation

```{r}
full_model <- function(test.id, myData, X, Y) {
  # full model
  full.model = lm(Y ~ ., data = myData[-test.id,])
  Ytest.pred = predict(full.model, newdata = myData[test.id,])
  # MSPE[1] = mean((myData$Y[test.id] - Ytest.pred)^2)
  mean_out= mean((myData$Y[test.id] - Ytest.pred)^2)
  return (mean_out)
}
```

# output variables for plotting

```{r}
final_output_df_bostonhousing2 = data.frame(Full=rep(0,50), R_min=rep(0,50),R_1se=rep(0,50), L_min=rep(0,50),L_1se=rep(0,50), L_Refit=rep(0,50),PCR=rep(0,50))
final_output_df_bostonhousing2tm = data.frame(Full=rep(0,50))
final_output_df_bostonhousing3 = data.frame(R_min=rep(0,50),R_1se=rep(0,50), L_min=rep(0,50),L_1se=rep(0,50), L_Refit=rep(0,50),PCR=rep(0,50))
```

# function to get bostondata2.csv MSE on 50 different training and test set predictions

```{r}
predict_bostondata2 <- function() {
  input_list = get_data("BostonData2.csv")
  myData = input_list$myData
  X = input_list$X
  Y = input_list$Y
  all.test.id = input_list$all.test.id
  
  # generate train and test
  # myData = read.csv("BostonData2.csv")
  # myData = myData[, -1]
  # X = data.matrix(myData[,-1])
  # Y = myData[,1]
  # T = 50
  # n = length(Y)
  # ntest = round(n * 0.25)  # test set size
  # ntrain = n - ntest  # training set size
  # all.test.id = matrix(0, ntest, T)  #
  # for(t in 1:T){
  #   all.test.id[, t] = sample(1:n, ntest)
  # }
  # 
  # print(all.test.id2)
  # print("-------------------------------------------------------------------------------------------")
  # print(all.test.id)
  for (m in 1:50) {
    test.id = all.test.id[,m] 
    # full.model = lm(Y ~ ., data = myData[-test.id,])
    # Ytest.pred = predict(full.model, newdata = myData[test.id,])
    # final_output_df_bostonhousing2tm[m,1] = mean((myData$Y[test.id] - Ytest.pred)^2)
    # print(test.id)
    # print(final_output_df_bostonhousing2tm)
    final_output_df_bostonhousing2tm[m,1]  = full_model(test.id, myData, X, Y)
  }
   return(final_output_df_bostonhousing2tm)
}
```


# main function to execute 50 different training and test simulations

```{r}
final_output_df_bostonhousing2tm = predict_bostondata2()
```

# chart bostondata2.csv MSE

```{r}
boxplot(final_output_df_bostonhousing2tm,main="MSPE for all modes after 50 simulations",
        ylab="Error",
        # col="orange",
        border="purple", las=2 )
```

