---
title: "Assignment_2_3102_bv10"
author: Bhuvaneswari(Bhu) Venkatraman
netid: bv10
seedid: 3102
output: html_document
---

#### netid: bv10

#### seedid: 3102

# Setting Seed and defining static variables


```{r}
library(glmnet) 
library(pls)
set.seed(3102)
```


# data read, training and test set generation as per dataset file name

```{r}
get_data <- function(dataset_file_name)  {
  
  # load csv
  myData = read.csv(dataset_file_name)
  myData = myData[, -1]
  X = data.matrix(myData[,-1])  
  Y = myData[,1] 
  
  # # generate train and test
  T = 50
  n = length(Y)
  ntest = round(n * 0.25)  # test set size
  ntrain = n - ntest  # training set size
  all.test.id = matrix(0, ntest, T)  #
  for(t in 1:T){
    all.test.id[, t] = sample(1:n, ntest)
  }
  
  #return necessary inputs to generate prediction
  return(list(myData=myData, X=X, Y=Y, all.test.id=all.test.id))
}
```

# full model prediction and MSE calculation

```{r}
full_model <- function(test.id, myData, X, Y) {
  # full model
  full.model = lm(Y ~ ., data = myData[-test.id,])
  Ytest.pred = predict(full.model, newdata = myData[test.id,])
  mean_out= mean((myData$Y[test.id] - Ytest.pred)^2)
  return (mean_out)
}
```


# ridge min, 1se

```{r}
ridge_model <- function(test.id, myData, X, Y) {
  # ridge model
  mean_out=rep(0,2)
  cv.out = cv.glmnet(X[-test.id, ], Y[-test.id], alpha = 0)
  best.lam = cv.out$lambda.min
  Ytest.pred = predict(cv.out, s = best.lam, newx = X[test.id, ])
  
  mylasso.lambda.seq = exp(seq(-10, -2, length.out = 100))
  cv.out = cv.glmnet(X[-test.id, ], Y[-test.id], alpha = 0, 
                     lambda = mylasso.lambda.seq)
  best.lam = cv.out$lambda.min
  Ytest.pred = predict(cv.out, s = best.lam, newx = X[test.id, ])
  mean_out[1] = mean((Y[test.id] - Ytest.pred)^2)
  
  best.lam = cv.out$lambda.1se
  Ytest.pred = predict(cv.out, s = best.lam, newx = X[test.id, ])
  mean_out[2] = mean((Y[test.id] - Ytest.pred)^2)
  return (mean_out)
}
```


# lasso min, 1se, refit

```{r}
lasso_model <- function(test.id, myData, X, Y) {
  # lasso model
  mean_out = rep(0,3)
  cv.out = cv.glmnet(X[-test.id, ], Y[-test.id], alpha = 1)
  best.lam = cv.out$lambda.min
  Ytest.pred = predict(cv.out, s = best.lam, newx = X[test.id, ])
  mean_out[1] = mean((Y[test.id] - Ytest.pred)^2)
  
  best.lam = cv.out$lambda.1se
  Ytest.pred = predict(cv.out, s = best.lam, newx = X[test.id, ])
  mean_out[2] = mean((Y[test.id] - Ytest.pred)^2)
  
  mylasso.coef = predict(cv.out, s = best.lam, type = "coefficients")
  var.sel = row.names(mylasso.coef)[which(mylasso.coef != 0)[-1]]
  mylasso.refit = lm(Y ~ ., myData[-test.id, c("Y", var.sel)])
  Ytest.pred = predict(mylasso.refit, newdata = myData[test.id, ])
  mean_out[3] = mean((Ytest.pred - Y[test.id])^2)
  
  return (mean_out)
}
```


# pcr

```{r}
pcr_model <- function(test.id, myData, X, Y) {
  # pcr model
  mypcr = pcr(Y ~ ., data= myData[-test.id, ], validation="CV")
  CVerr = RMSEP(mypcr)$val[1, , ]
  adjCVerr = RMSEP(mypcr)$val[2, , ]
  best.ncomp = which.min(CVerr) - 1 
  if (best.ncomp==0) {
    Ytest.pred = mean(myData$Y[-test.id])
  } else {
    Ytest.pred = predict(mypcr, myData[test.id,], ncomp=best.ncomp)
  }
  
  mean_out = mean((Ytest.pred - myData$Y[test.id])^2)
  return (mean_out)
}
```


# output variables for plotting

```{r}
# final_output_df_bostonhousing2 = data.frame(Full=rep(0,50), R_min=rep(0,50),R_1se=rep(0,50), L_min=rep(0,50),L_1se=rep(0,50), L_Refit=rep(0,50),PCR=rep(0,50))
final_output_df_bostonhousing3 = data.frame(R_min=rep(0,50),R_1se=rep(0,50), L_min=rep(0,50),L_1se=rep(0,50), L_Refit=rep(0,50),PCR=rep(0,50))
```

# function to get bostondata3.csv MSE on 50 different training and test set predictions

```{r}
predict_bostondata3 <- function() {
  input_list = get_data("BostonData3.csv")
  myData = input_list$myData
  X = input_list$X
  Y = input_list$Y
  all.test.id = input_list$all.test.id
  
  for (m in 1:50) {
    test.id = all.test.id[,m] 
    ridge_out = ridge_model(test.id, myData, X, Y)
    final_output_df_bostonhousing3[m,1]  = ridge_out[1]
    final_output_df_bostonhousing3[m,2]  = ridge_out[2]
    lasso_out = lasso_model(test.id, myData, X, Y)
    final_output_df_bostonhousing3[m,3]  = lasso_out[1]
    final_output_df_bostonhousing3[m,4]  = lasso_out[2]
    final_output_df_bostonhousing3[m,5]  = lasso_out[3]
    final_output_df_bostonhousing3[m,6]  = pcr_model(test.id,myData, X, Y)
  }
  return(final_output_df_bostonhousing3)
}
```

# main function to execute 50 different training and test simulations

```{r}
final_output_df_bostonhousing3 = predict_bostondata3()
```

# chart bostondata3,,,,,,,,rfmhko .csv MSE

```{r}
final_output_df_bostonhousing3
boxplot(final_output_df_bostonhousing3,main="MSPE for all modes after 50 simulations",
        ylab="Error",
        # col="orange",
        border="purple", las=2 )
```

```{r}
myDf1 <- as.data.frame(final_output_df_bostonhousing3)
myDf1
```

```{r}
data <- myDf1[c("L_Refit","L_min"  , "L_1se" , "R_min" ,  "R_1se", "PCR" )]
data

boxplot(data , col=rgb(0.3,0.5,0.4,0.6) , ylab="value" , 
    xlab="names in desired order")
```

